Adaptive Waveform Inversion for Transmitted Wave Data

### 2024-11-03 repro/papers/awi/paper.tex
rearrange

\title{Adaptive Waveform Inversion for Transmitted Wavefields}
\author{William W. Symes}

@@

\section{Introduction}
Full waveform inversion (FWI) refers to the inference of mechanical
parameter distribution within a material object  (the earth, a human
body, a manufactured artifact,...) from remote measurements of waves
propagating through it. Over the last several decades, this task has come to
be carried out by minimizing a measure of the difference between computational
prediction of wave data versus actual observed
data. The mechanical parameters to be inferred appear as
coefficients in systems of partial differential equations governing
the wave motion, and discrete representations of these parameters form
the unknown vectors in the optimization problems thus generated.

While widely used in recent years, FWI exhibits a pathology when wave
velocities or their parametric equivalents are amongst the parameters
to be estimated: the excessive sensitivity of predicted waveforms to
changes in velocity leads to oscillations in the difference measures,
especially for the most commonly used square-integral measure. Local
optimization algorithms (relatives of Newton;'s method) are the only practical
methods to generate sequences of approximate optimal
parameters. However the oscillations in the difference measures lead
these sequences to stagnate at parameter distributions
far from the optimum, and thus do not extract the maximum information
from the data \cite[]{VirieuxOperto:09}.

Various modifications of straighforward best-fit FWI have been
suggested to overcome this ``cycle-skipping'' pathology. Many of these
modifications employ {\em model extension}, that is, the introduction
of a larger space of model parameters than derives from basic
physics. An example is the {\em source-receiver} extension \cite[]{HuangSymes:17}, in which
source waveforms are permitted to depend (unphysically) on receiver
locations. The focus of this paper is Adaptive Waveform Inversion
\cite[]{Warner:16} (AWI), a modification of FWI using the
source-receiver extension. AWI has been successfully applied to industry-scale seismic data
\cite[]{GuaschWarnerRavaut:GEO19,Warneretal:SEG21} and shows promise for
ultrasonic imaging of the the human brain
\cite[]{Guaschetal:NPJDM20}.

While these and other works discuss heuristic justifications for AWI
and many details of implementation, a clear explanation for its
effectiveness is so far missing from the literature. In this short
note, I treat a special case of FWI, the recovery of smooth acoustic
medium parameters from data generated at localized energy sources,
propagated through a fluid, and recorded at receiver locations remote
from the sources. I show that the AWI objective function is an
asymptotic (in time frequency) approximation to the mean-square
time-of-travel misfit, provided that arrival times of transmitted
waves are single, that is, unique rays of geometric asymptotics
connect sources and receivers. The error in this approximation is
proportional to the source mean-square wavelength. Thus AWI and
minimization of mean-square travel time error (travel time tomography)
lead to similar material parameter estimates. Travel time tomography
appears to be much less sensitive to choice of initial model than is
FWI, and in fact is commonly used to construct an initial model for
FWI \cite[]{Bordingetal:87,SirguePratt:04,VirieuxOperto:09}. At least
in the special case treated here, AWI implicitly combines travel time
tomography with FWI.

The reader will
note that a normalization component of the AWI objective function
definition is important in reaching this conclusion. The connection
between objective normalization and the asymptotic equivalence to
travel time inversion for transmission data has nothing to do with the
reasons for normalizing the objective previoiusly advanced in the
literature \cite[]{Warner:16,LiAlkhalifah:21}. Another modification of
FWI based on source-receiver extension, Matched Source Waveform Inversion (MSWI)
\cite[]{HuangSymes2015SEG,HuangSymes:Geo17} lacks the AWI
normalization, and has a looser relation to travel time inversion. The
arguments presented here for AWI (and MSWI) are refinements of those
given by \cite{HuangSymes:Geo17}. which were in turn based on ideas
presented in Song's thesis \cite[]{Song:94c}.

After establishing the class of problems to be studied and notations
for their description in the following section, I present the
asymptotic analysis leading to the main conclusion in the next. In
effect, AWI regularizes or relaxes the FWI problem. For many inverse
problems, regularization has been accomplished via additive penalty
functions. In section four, I explain the relation between AWI and a
penalty formulation of the wave inverse problem. The close relation
between AWI objective and travel time misfit is limited in scope: in
the fifth section, I show that it does not hold in general for
transmitted wave data in which multiple rays connect source and
receiver.

@@

\section{Adaptive Waveform Inversion}

The version of AWI introduced by \cite{Warner:16} is based on an
active source acoustic model of data generation in Euclidean space
$\bR^3$. In this section, I will review a version of this model, place
some useful constraints on energy sources appearing in it, and define
AWI as an optimization problem.

\subsection{Acoustic Modeling of Transmitted Waves}

@@

The pressure and velocity fields $p({\bf x},t;{\bf x}_s)$,
${\bf v}({\bf x},t;{\bf x}_s)$ for the source location ${\bf x}_s$
depend on the bulk modulus $\kappa$, buoyancy $\beta$
(reciprocal of the density $\rho$), and source wave
form $f(t)$ through the acoustic system
\begin{eqnarray}
  \label{eqn:awe}
 \frac{\partial p}{\partial t} & = &- \kappa \nabla \cdot {\bf v} +
                                    f(t) \delta({\bf x}-{\bf x}_s); \nonumber \\
\frac{\partial {\bf v}}{\partial t} & = & - \beta \nabla p; \\ 
p, {\bf v} & = & 0 \mbox{ for }  t \ll 0.
\end{eqnarray}
For simplicity, I assume that the system \ref{eqn:awe} holds over all of Euclidean
space-time $(\bx,t) \in \bR^3 \times \bR$, and ignore physical
boundaries such as the Earth's surface. Likewise for simplicity, I
assume in this study that the material density $\rho$ is spatially
homogeneous and known. In some settings, eg. the earth's
upper crust, the material density $\rho$ exhibits substantially less variance than the
bulk modulus, Density is however spatially variable in most geological
and biomedical materials, and could be included amongst the parameters
to be estimated.

The parameter to be determined in the inverse problem studied here is
the bulk modulus $\kappa$. It is presumed is bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Under these hypotheses, the symmetric hyperbolic system \ref{eqn:awe}
has a unique distribution
solution if $f \in \cal{E}'(\bR)$ (for example \cite{Lax:PDENotes},
Theorem 7.4). The Propagation of
Singularities Theorem (\cite{Tay:81}, Theorem VI.2.1) and
the pull-back property of distributions (\cite{Dui:95}, Proposition 1.3.3)
imply that if $f \in \cal{E}'(\bR)$ and $(\bx_s,\bx_r) \in X_{sr}$,
then the trace $p(\bx_r, \cdot;\bx_s)$ on $\{(\bx_r,t): t \in \bR\}$ is well-defined,
as a distribution vanishing for sufficiently large
negative $t$. Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
  \rightarrow X_{sr} \times {\cal D}'(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa]= p|_{X_{sr} \times \bR}
\end{equation}

In this setting, acoustic FWI means: given the set of
source-receiver position pairs $X_{sr}$, source wavelet
$f \in {\cal E}'(\bR)$,  density $\rho$, and data traces
$d:  X_{sr} \rightarrow \cal{D}(\bR)$, find a bulk modulus $\kappa$
(or equivalently wave velocity $c=\sqrt{\kappa/\rho}$)
so that $F[\kappa] \approx d$ in some sense. If
$d$ and $F[\kappa]$ lie in a normed subspace of ${\cal D}'(\bR)$ for each
$(\bx_s,\bx_r) \in X_{sr}$, then the
norm provides a straightforward candidate for this sense of
approximation.

%In practice, recording takes place over a finite time interval
%$[0,t_d]$. For convenience, assume that this interval is the same for
%all receivers. Let $\phi \in C_0^{\infty}(0,t_d)$ be a window function
%supported in this interval. Additional constraints on $t_d$ and $\phi$
%will be be specified below.

@@

The arguments to follow depend an additional assumption about solutions of
the wave equation \ref{eqn:awe2}, restricted to $X_{sr}\times \bR$:
\begin{itemize}
\item[G4. ] For any $n > 0$, $f \in {\cal E}'(\bR)$, there exist $K >
  0, T \in \bR,$ and $\delta>0$ so
  that for any $k$ with $0 \le k \le n$, $(\bx_s,\bx_r) \in
 X_{sr},$
  and any $\log \kappa \in M$, the solution $p$ of \ref{eqn:awe2} is smooth
  for $t>T$ and satisfies
  \[
    |\partial_t^k p(\bx_r,t;\bx_s)| \le K e^{-\delta t}, \, t > T.
  \]
\end{itemize}
Condition G4 is a special case of {\em local energy decay}, a much
studied topic. Properties which imply G4 are in turn implied by the
{\em non-trapping} hypothesis on the model $\log \kappa \in M$: all rays of
geometric asymptotics escape to infinity (\cite{Hristova:09}, Theorem
2, \cite{EgorovShubin}, Theorem 2.104, p. 192).

@@

\subsection{Notational Conventions for Norms and Bounds}
The Banach spaces
$l^p(X_{sr},L^p(\bR))$ will be abbreviated as $L^p(X_{sr} \times
\bR)$. The norm in a Banach space $B$ will be denoted
$\|\cdot\|_{B}$. The norm symbol
$\|\cdot\|$ without subscript will be understood to mean either
$\|\cdot\|_{L^2(\bR)}$ or $\|\cdot\|_{L^2(X_{sr}\times
  \bR)}$, as context indicates.

\subsection{FWI via Least Squares}
For $f \in {\cal E}'(\bR)$, the Hadamard decomposition 
(equation \ref{eqn:regreen}) gives:
\begin{equation}
  \label{eqn:fwdgreen}
  F[\kappa] = G[\kappa] * \rho \partial_t f  = \rho( a[\kappa]\delta
  +  (b[\kappa]H)) *\delta_{\tau[\kappa]}*\partial_t f.
\end{equation}
Suppose that the source wavelet $f$ satisfies
\begin{itemize}
\item[F1. ] $f \in C^{\infty}_0(\bR)$;
\item[F2. ] $f$ is odd, $f(t) = -f(-t)$.
\end{itemize}
Then equation \ref{eqn:fwdgreen}, assumptions F1, F2, bounds
\ref{eqn:best} and Young's inequality implies that $F[\kappa] \in
L^p(X_{sr} \times \bR)$ for any $p \ge 1$. Moreover there exists
$C>0$ so that for $\log \kappa\in M$, $p \ge 1$,
\begin{equation}
  \label{eqn:fwdnorm}
  \|F[\kappa]\|_{L^p(X_{sr} \times \bR)} \le C \left\|\partial_t f\right\|_{L^p(\bR)}.
\end{equation}
This observation justfies posing FWI as a least-squares problem: given $d \in
L^2(X_{sr}\times \bR)$, $f \in H^1_0(\bR)$, and $\rho_s>0$, find $\kappa$ to minimize
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[m;d]= \frac{1}{2}\|F[\kappa]-d\|_{L^2(X_{sr} \times \bR)},
\end{equation}

@@

\subsection{Tihonov Regularization}
Before establishing the relation with travel time tomography, several
defects in the definition \ref{eqn:awideforig} of $J_{\rm AWI}$ must be
remedied. First, the objective function is not well-defined for the
``ideal'' case, in which $d = F[\kappa]$ and $u[\kappa;d]=\delta$:
this adaptive filter is not square-integrable, so the denominator does
not make sense. Second, the equation $S[\kappa]u=d$ does not (in
general) have a solution $u \in L^2(X_{sr} \times \bR)$. Since
$S[\kappa]$ is a convolution operator, solving $S[\kappa]u=d$ for $u$
amounts to deconovolving $d$ by $F[\kappa]$..  Under the conditions
W1-W3 on the wavelet $w$, $F[\kappa]$ does not have a continuous
(stable) convolution inverse, and $S[\kappa]$ is not surjective.

In fact, these two defects are really symptoms of the instability of
deconvolution for kernels with the properties of $F[\kappa]$.  From \ref{eqn:fwdgreen} and the
definition \ref{eqn:sdef} of  $S[\kappa]$,
\[
  S[\kappa]u = \rho( a[\kappa]\delta    +  (b[\kappa]H)) *\delta_{\tau[\kappa]}*\partial_t f *u,
\]
\begin{equation}
  \label{eqn:sgood}
  = \rho(a[\kappa]\delta_{\tau[\kappa]}*\partial_t f +
  (b[\kappa](0)\delta + (\partial_t b[\kappa])H)
  *\delta_{\tau[\kappa]}*f) * u.
\end{equation}
Since $\partial_t f$ is even, the formal adjoint $S[\kappa]^T$ is given by
\begin{equation}
  \label{eqn:sadjgood}
  S[\kappa]^Tu = \rho( a[\kappa]\delta    +  (\check{b}[\kappa]\check{H}))
  *\delta_{-\tau[\kappa]}*\partial_t f *u,
\end{equation}
in which $\check{u}(t) = u(-t)$.
The key properties of $S$ follow directly from Young's inequality and
the representation \ref{eqn:sgood}:
\begin{lemma}
  \label{thm:snorm}
  Under the assumptions G1-G4 leading to \ref{eqn:fwdgreen} and
  the properties F1 and F2, the expressions \ref{eqn:sgood} and \ref{eqn:sadjgood} define 
  bounded operators on
  $\Pi_{X_{sr}} {\cal S}(\bR)$ and on
  $L^2(X_{sr} \times \bR)$, and there exists $C>0$ so
  that for any $\kappa \in M$,
  \begin{equation}
    \label{eqn:snorm}
    \|S[\kappa]\| \le C\|\partial_t f\|_{L^1(\bR)}.
  \end{equation}
\end{lemma}
\begin{proof} From the estimates H3, $b[\kappa](\bx_r,\cdot;\bx_s)H *
  \partial_tf \in {\cal S}(\bR)$ for all $(\bx_s,\bx_r) \in X_{sr}$,
  whence the first conclusion. Bounded extension to $L^2(X_{sr} \times
  \bR)$ follows from Young's inequality.
\end{proof}

>>>>

\title{Adaptive Waveform Inversion for Transmitted Wave Data}
\author{William W. Symes}

@@

\section{Introduction}
Full waveform inversion (FWI) refers to the inference of mechanical
parameter distribution within a material object  (the earth, a human
body, a manufactured artifact,...) from remote measurements of waves
propagating through it. These mechanical parameters appear as
coefficients in systems of partial differential equations governing
wave propagation, and numerical solutions of these wave equations can
be used to predict experimental data {\em ab initio}. FWI consists in
updating the parameters to minimize a numerical measure of  the difference between computationally
predicted wave data and actual observed
data. If wave velocities or their parametric equivalents are amongst the parameters
to be estimated, such numerical measures tend to oscillate, due to:
the excessive sensitivity of predicted waveforms to changes in wave velocity.
%especially for the most commonly used square-integral measure ($L^2$
%norm).
These oscillations cause iterative local optimization algorithms (relatives of
Newton;'s method, the only practical algorithmic choices) to stagnate
far from good data fit \cite[]{VirieuxOperto:09}, unless the initial
parameter estimates are of very high quality.

Adaptive Waveform
Inversion \cite[]{Warner:16} (AWI) is a modification of FWI, designed to overcome this
``cycle-skipping'' pathology. AWI has been successfully applied to
industry-scale seismic data
\cite[]{GuaschWarnerRavaut:GEO19,Warneretal:SEG21} and shows promise
for ultrasonic imaging of the the human brain
\cite[]{Guaschetal:NPJDM20}. These and other works discuss heuristic justifications for AWI
and many details of implementation. However, a clear explanation for its
effectiveness is so far missing from the literature.

This paper supplies such an explanation for a special case of AWI, the recovery of smooth acoustic
medium parameters from data generated at localized energy sources,
propagated through a fluid, and recorded at receiver locations remote
from the sources. I show that the AWI objective function is an
asymptotic (in time frequency) approximation to the mean-square
time-of-travel misfit, provided that arrival times of transmitted
waves are single, that is, unique rays of geometric asymptotics
connect sources and receivers. The error in this approximation is
proportional to the source mean-square wavelength. Thus AWI and
minimization of mean-square travel time error (travel time tomography)
lead to similar material parameter estimates. Travel time tomography
appears to be much less sensitive to choice of initial model than is
FWI, and in fact is commonly used to construct an initial model for
FWI \cite[]{Bordingetal:87,SirguePratt:04,VirieuxOperto:09}. At least
in the special case treated here, AWI implicitly combines travel time
tomography with FWI.

The reader will
note that a normalization component of the AWI objective function
definition is important in reaching this conclusion. The connection
between objective normalization and the asymptotic equivalence to
travel time inversion for transmission data has nothing to do with the
reasons for normalizing the objective previoiusly advanced in the
literature \cite[]{Warner:16,LiAlkhalifah:21}. Another similar modification of
FWI, Matched Source Waveform Inversion (MSWI)
\cite[]{HuangSymes2015SEG,HuangSymes:Geo17} lacks the AWI
normalization, and has a looser relation to travel time inversion. The
arguments presented here for AWI (and MSWI) are refinements of those
given by \cite{HuangSymes:Geo17}. which were in turn based on ideas
presented in Song's thesis \cite[]{Song:94c}.

The next section establishes the class of problems to be studied and notations
for their description, defines the AWI and MSWI objective functions,
and states the main results of this paper. in the following section, I present the
asymptotic analysis leading to the main results. In
effect, AWI regularizes or relaxes the FWI problem. For many inverse
problems, regularization has been accomplished via additive penalty
functions. In section four, I explain the relation between AWI and a
penalty formulation of the wave inverse problem. The close relation
between AWI objective and travel time misfit is limited in scope: in
the fifth section, I show that it does not hold in general for
transmitted wave data in which multiple rays connect source and
receiver.

@@

\section{Main Results}

The version of AWI introduced by \cite{Warner:16} is based on an
active source acoustic model of data generation in Euclidean space
$\bR^3$.

@@

Throughtout the following discussion, it is assumed that
$X_{sr}$ is finite.

The pressure and velocity fields $p({\bf x},t;{\bf x}_s)$,
${\bf v}({\bf x},t;{\bf x}_s)$ for the source location ${\bf x}_s$
depend on the bulk modulus $\kappa$, buoyancy $\beta$
(reciprocal of the density $\rho$), and source wave
form $f \in C_0^{\infty}(\bR)$ through the acoustic system
\begin{eqnarray}
  \label{eqn:awe}
 \frac{\partial p}{\partial t} & = &- \kappa \nabla \cdot {\bf v} +
                                    f(t) \delta({\bf x}-{\bf x}_s); \nonumber \\
\frac{\partial {\bf v}}{\partial t} & = & - \beta \nabla p; \\ 
p, {\bf v} & = & 0 \mbox{ for }  t \ll 0.
\end{eqnarray}
For simplicity, I assume that the system \ref{eqn:awe} holds over all of Euclidean
space-time $(\bx,t) \in \bR^3 \times \bR$, and ignore physical
boundaries such as the Earth's surface. Likewise for simplicity, I
assume in this study that the material density $\rho$ is spatially
homogeneous and known.
%In some settings, eg. the earth's
%upper crust, the material density $\rho$ exhibits substantially less variance than the
%bulk modulus, Density is however spatially variable in most geological
%and biomedical materials, and could reasonabliy be included amongst the parameters
%to be estimated.

The parameter to be determined in the inverse problem studied here is
the bulk modulus $\kappa$. It is presumed is bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Under these hypotheses, the symmetric hyperbolic system \ref{eqn:awe}
has a unique distribution solution (for example \cite{Lax:PDENotes},
Theorem 7.4). The Propagation of
Singularities Theorem (\cite{Tay:81}, Theorem VI.2.1) %and
%the pull-back property of distributions (\cite{Dui:95}, Proposition
%1.3.3)
imply that $p(\cdot,\cdot;\bx_s)$ is smooth away from $\{\bx_s\}
\times \bR$, so that the traces $p(\bx_r, \cdot;\bx_s)$ on
$\{(\bx_r,t): t \in \bR\}$ are well-defined.

Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: exp(M)  \rightarrow \Pi_{X_{sr}} {\cal D}'(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa]= p|_{X_{sr} \times \bR}
\end{equation}

In this setting, acoustic FWI means: given the set of
source-receiver position pairs $X_{sr}$, source wavelet
$f \in C_0^{\infty}(\bR)$,  density $\rho$, and data traces
$d \in \Pi_{X_{sr}} C^{\infty}(\bR)$, find a bulk modulus $\log
\kappa$
(or equivalently wave velocity $c=\sqrt{\kappa/\rho}$)
so that $F[\kappa] \approx d$ in some sense. If
$d(\bx_r,\cdot;\bx_s)$ and $F[\kappa](\bx_r,\cdot;\bx_s)$ lie in a normed subspace of ${\cal D}'(\bR)$ for each
$(\bx_s,\bx_r) \in X_{sr}$, then the
norm provides a straightforward candidate for this sense of
approximation.

The analysis to follow requires two further constraints on the domain
$M \subset C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)$ of $F$. The first is
the {\em Exponential Decay} assumption:
\begin{quote}
  For any $n > 0$, $f \in {\cal E}'(\bR)$,
  there exist $K >
  0, T \in \bR,$ and $\delta>0$ so
  that for any $k$ with $0 \le k \le n$, $(\bx_s,\bx_r) \in
 X_{sr},$ and any $\log \kappa \in M$, the solution $(p,\bv)$ of \ref{eqn:awe} is smooth
  for $t>T$ and satisfies
  \[
    |\partial_t^k p(\bx_r,t;\bx_s)| \le K e^{-\delta t}, \, t > T.
  \]
\end{quote}
SInce $X_{sr}$ is bounded, the property described in this assumption
is a special case of {\em local energy decay}, a much studied topic. The {\em
  non-trapping} hypothesis is sufficient to ensure local energy decay:
all rays of geometric asymptotics (that is, geodesics for the Riemannian metric associated
to the system \ref{eqn:awe}) escape to infinity (\cite{Hristova:09},
Theorem 2, \cite{EgorovShubin}, Theorem 2.104, p. 192). That is,
exponential decay follows from a characteristic of $\kappa$.

The second assumption explicitly involves a constraint on ray geometry
(therefore on $\kappa$). A {\em normal neighborhood} of
$\bx_s \in \bR^3$ is an open subset $\Omega$ of $\bR^3$ 
within which every point is connected to $\bx_s$ by  a unique ray  (\cite{Friedlander:75},
p. 16). Therefore, to every point $\bx \in \Omega$ is associated a
unique time of travel (geodesic distance) $\tau(\bx_s,\bx)$ from $\bx_s$ to $\bx$.
The {\em Single Arrival} assumption is
\begin{quote}
  $\Omega\subset \bR^3$ 
and $M \subset C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)$ are bounded open sets satisfying 
\begin{itemize}
  \item[G1. ] $X_{sr} \subset \Omega \times \Omega$;
  \item[G2. ] for every source point $\bx_s\in X_{s}$ and $\log \kappa \in M$, $\Omega$ is a
    normal neighborhood of $\bx_s$.
  \item[G3. ] for every source point $\bx_s \in X_{s};$
    $\log \kappa \in M$, ray ${\bf X}:\bR^+\rightarrow \bR^3$
    with ${\bf X}(0)=\bx_s$ and any $t>0$, either $\{{\bf X}(s): 0\le s \le t\} \subset \Omega$
    or $\{{\bf X}(s): t < s < \infty\} \subset \bR^3 \setminus \Omega$.
  \end{itemize}
\end{quote}
Assertion G3 is the {\em no-return} assumption: once a ray leaves
$\Omega$, its subsequent intersection with $\Omega$ is null.

Thanks to exponential decay, each trace $p(\bx_r,\cdot;\bx_s)$ is in $L^{p}(\bR)$ for $p
\ge 1$.
Denote by $L^p(X_{sr} \times \bR)$ the product space $l^p(X_{sr},L^p(\bR))$
with the usuall product norm.

Accordingly, FWI may be posed as a least-squares
problem: given data $d \in
L^2(X_{sr}\times \bR)$, find $\kappa$ to minimize
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[\kappa;d]= \frac{1}{2}\|F[\kappa]-d\|^2_{L^2(X_{sr} \times \bR)},
\end{equation}
As mentioned earlier, optimization algorithms related to Newton's method, applied to
$J_{\rm FWI}$, tend to stagnate far from global minimizers if
the initial choice of $\kappa$ does not predict the arrival times of
waves correctly within a half-wavelength or so (the meaning of
``wavelength'' in this context will be explained below)
\cite[]{GauTarVir:86,VirieuxOperto:09}.

AWI modifies the least-squares FWI problem by interposing a convolution operator or
{\em filter} $u \in L^2(X_{sr} \times \bR)$. That is, the residual
data $F[\kappa]-d$ is replaced by the filtered residual data $F[m]*u-d$.
Since $F[\kappa] \in L^1(X_{sr}\times \bR)$, the filtererd residual
makes sense.
The filter $u$ is intended to shift localized wave energy in $F[m]$ to
match that present in $d$. If on the other hand $u(\bx_r,t;\bx_s) =
\delta(t)$ for each $(\bx_x,\bx_r) \in X_{sr}$, then $tu = 0$. This
observation motivates the definition of the AWI objective function
\begin{equation}
  \label{eqn:awidefmain}
  J_{\rm AWI,\sigma}[\kappa;d] = \sum_{(\bx_s,\bx_r) \in X_{sr}}
  \frac{\|Tu_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2}{\|u_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2}.
\end{equation}
in which $Tu(\bx_r,t;\bx_s) = tu(\bx_r,t;\bx_s)$ and
\begin{equation}
  \label{eqn:reg}
  u_{\sigma}[\kappa;d] = \mbox{argmin}_{u \in L^2(X_{sr} \times \bR)} (\|F[\kappa]*u-d\|^2 + \sigma\|u\|^2)
\end{equation}
The objective function for the closely related MSWI approach,
mentioned in the introduction, is  
\begin{equation}
  \label{eqn:mswidefmain}
  J_{\rm MSWI,\sigma}[\kappa;d] = \sum_{(\bx_s,\bx_r) \in X_{sr}}
  \|Tu_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2.
\end{equation}
Note that while the Lax-Milgram Theorem (\cite{Yosida}, section
III.7) ensures that $u_{\sigma}[\kappa;d] \in L^2(X_{sr}\times \bR)$,
it is not immediiately obvious that $Tu_{\sigma}[\kappa;d] \in L^2(X_{sr}\times \bR)$.

Asymptotic analysis of AWI and MSWI is based on a family of source
wavelet: for $\lambda > 0$,
\begin{equation}
  \fl(t)= \lambda^{\frac{1}{2}}f_1\left(\frac{t}{\lambda}\right).
\end{equation}
The relation between the dimensionless scale parameter $\lambda$ and
RMS wavelength will be explained below. Correspondingly, denote the
modeling operator with $f=\fl$ by $\Fl$.

The main result of this analysis is

\begin{theorem}
  \label{thm:main}
  Assume that the Exponential Decay and Single Arrival assumptions
  hold for $X_{sr} \subset \Omega \times \Omega, M \subset C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)$.
  Select $r >0$ with units of
  pressure$^2 \times$time$^2$ and $\log \kappa^* \in M$. For $0 <
  \lambda $, set
  $\dl=\Fl[\kappa^*]$ (``noise-free data''), and  define $\ul[\kappa,\dl]$
  to be the solution of 
  equation \ref{eqn:reg} with  $\sigma=r\lambda$, $F=\Fl$ and $d=\dl$..
  There exist $0 < C_* \le C^*$, $0 < C$, and $W: M \times X_{sr}
  \rightarrow \bR^+$ with $C_* \le
  W \le C^*$ so that for
  $\log \kappa \in M$, $(\bx_s,\bx_r) \in X_{sr}$, $Tu_{\sigma}[\kappa;\dl](\bx_r,\cdot;\bx_s) \in L^2(\bR)$, and 
  \begin{equation}
    \label{eqn:mswitt}
    |\|T\ul[\kappa;\dl](\bx_r,\cdot;\bx_s)\|- \lambda^{-\frac{1}{2}} (W[\kappa^*,\kappa](\bx_s,\bx_r))^{\frac{1}{2}}|\tau[\kappa](\bx_s,\bx_r)-\tau[\kappa^*](\bx_s,\bx_r)|| \le C
  \lambda^{\frac{1}{2}}
\end{equation}
and for sufficiently small $\lambda$, 
\begin{equation}
  \label{eqn:awitt}
  \left|\frac{\|T\ul[\kappa;\dl](\bx_r,\cdot;\bx_s)\|}{\|\ul[\kappa,\dl](\bx_r,\cdot;\bx_s)\|}-
    |\tau[\kappa](\bx_s,\bx_r)-\tau[\kappa^*](\bx_s,\bx_r)|\right| \le C \lambda
  \end{equation}
\end{theorem}

Summation over the source-receiver positions in $X_{sr}$ yields
\begin{cor}
  \label{thm:cor1}
  In addition to the assumptions of Theorem \ref{thm:main}, define $J_{\rm
    AWI,\sigma}$ and $J_{\rm MSWI,\sigma}$ by equations
  \ref{eqn:awidefmain}, \ref{eqn:mswidefmain} respectively. Then
\begin{equation}
\label{eqn:cor1mswi}   
\lim_{\lambda \rightarrow 0} \lambda J_{\rm MSWI,r\lambda}[\kappa;\dl]
= \sum_{(\bx_s,\bx_r) \in X_{sr}} W[\kappa^*,\kappa^*]^2(\bx_s,\bx_r) |\tau[m](\bx_s,\bx_r) - \tau[m^*](\bx_s,\bx_r)|^2.
\end{equation}
and
\begin{equation}
\label{eqn:cor1awi}   
\lim_{\lambda \rightarrow 0} J_{\rm AWI,r\lambda}[\kappa;\dl]
= \sum_{(\bx_s,\bx_r) \in X_{sr}}  |\tau[m](\bx_s,\bx_r) - \tau[m^*](\bx_s,\bx_r)|^2.
\end{equation}
\end{cor}

@@

@@

\subsection{FWI via Least Squares}
For $f \in {\cal E}'(\bR)$, the Hadamard decomposition 
(equation \ref{eqn:regreen}) gives:
\begin{equation}
  \label{eqn:fwdgreen}
  F[\kappa] = G[\kappa] * \rho \partial_t f  = \rho( a[\kappa]\delta
  +  (b[\kappa]H)) *\delta_{\tau[\kappa]}*\partial_t f.
\end{equation}

For the remainder of this paper, source wavelet $f$ will be assumed
smooth, of compact support: $f \in C^{\infty}_0(\bR)$. Less stringent constraints on $f$ would
support the results below, but this is the simplest choice. Also, with
a few notable exceptions \cite[]{SondRes:83}, source wavelets that
adequately model physical experimentss may be take to be smooth of
compact support, because the physical phenomena that inject energy
into the acoustic field are transient, and the data is sampled at a
finite rate and either analog
or digitally filtered (or both) to well below Nyquist frequency.

Then equation \ref{eqn:fwdgreen}, bounds
\ref{eqn:best} and Young's inequality implies that $F[\kappa] \in
L^p(X_{sr} \times \bR)$ for any $p \ge 1$. Moreover there exists
$C>0$ so that for $\log \kappa\in M$, $p \ge 1$,
\begin{equation}
  \label{eqn:fwdnorm}
  \|F[\kappa]\|_{L^p(X_{sr} \times \bR)} \le C \left\|\partial_t f\right\|_{L^p(\bR)}.
\end{equation}
This observation justfies posing FWI as a least-squares problem: given $d \in
L^2(X_{sr}\times \bR)$, $f \in H^1_0(\bR)$, find $\kappa$ to minimize
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[m;d]= \frac{1}{2}\|F[\kappa]-d\|_{L^2(X_{sr} \times \bR)},
\end{equation}

@@

\subsection{A Note on Notations  for Norms and Bounds}
The Banach spaces
$l^p(X_{sr},L^p(\bR))$ will be abbreviated as $L^p(X_{sr} \times
\bR)$. The norm in a Banach space $B$ will be denoted
$\|\cdot\|_{B}$. The norm symbol
$\|\cdot\|$ without subscript will be understood to mean either
$\|\cdot\|_{L^2(\bR)}$ or $\|\cdot\|_{L^2(X_{sr}\times
  \bR)}$, as context indicates.

\subsection{Tihonov Regularization}
Before establishing the relation with travel time tomography, several
defects in the definition \ref{eqn:awideforig} of $J_{\rm AWI}$ must be
remedied. First, the objective function is not well-defined for the
``ideal'' case, in which $d = F[\kappa]$ and $u[\kappa;d]=\delta$:
this adaptive filter is not square-integrable, so the denominator does
not make sense. Second, the equation $S[\kappa]u=d$ does not (in
general) have a solution $u \in L^2(X_{sr} \times \bR)$. Since
$S[\kappa]$ is a convolution operator, solving $S[\kappa]u=d$ for $u$
amounts to deconovolving $d$ by $F[\kappa]$..  Under the standing
assumption that $f \in C^{\infty}_0(\bR)$, $F[\kappa]$ does not have a continuous
(stable) convolution inverse, and $S[\kappa]$ is not surjective.
